Relevance-based Quantization of Scattering Features for Unsupervised Mining of Environmental Audio

### 2018-08-17 paper/lostanlenRelevanceScattering.tex
add park_scalogram, include it in paper

The experiments in this paper are carried out on the DCASE 2013 dataset~\cite{7100934}.
Although the dataset was constructed for the task of acoustic scene classification, where the goal is to correctly assign the class of a given recording, we can use the same recordings and class labels for the task of acoustic scene similarity retrieval.
The dataset consists of two parts, a public and a private subset, each made up of $100$ acoustic scene recordings sampled at $44100\,\mathrm{Hz}$ and $30$ seconds in duration. The dataset is evenly divided into $10$ acoustic scene classes: \emph{bus}, \emph{busy street}, \emph{office}, \emph{open air market}, \emph{park}, \emph{quiet street}, \emph{restaurant}, \emph{supermarket}, \emph{tube}, and \emph{tube station}.
The recordings were made by three different recordists at a wide variety of locations in the Greater London area over a period of several months.
In order to avoid any correlation between recording conditions and label distribution, all recordings were carried out under moderate weather conditions, at varying times of day and week, and each recordist recorded each scene type. As a result, the dataset enjoys significant intra-class diversity while remaining of manageable size, making it suitable for evaluation of algorithmic design choices~\cite{lagrange:hal-01082501}.

@@

The scaling parameter $q$ of the RBF kernels (see Eq.~\ref{eq:kc}) is set to $10\%$ of the number of data points to cluster.  As the number of Gaussians for the BoF approach, the numbers of clusters $M$ controls the level of abstraction. For each method, the parameter $M$ is set to $8$. It thus allows $8$ different types of observations to de modeled, which seems reasonable given the duration of the scene (30 seconds). Note that this is the only free parameter in the proposed method. However, except for $\emph{RbQ-a}$, the results are not very sensitive to the choice of $M$, as long as it is large enough to characterize the seen. For a numerical demonstration of this, please see the end of the next section.

@@

\subsubsection*{Sensitivity to number of clusters $M$}

We now study the sensitivity of the precision at rank $5$ ($p@5$) with respect to the number of clusters $M$.
The results are shown in Figure \ref{fig:sensitivityM}.

For a small number of clusters ($M = 1$ or $M = 2$), all methods perform worse, since not enough discriminative sound objects are extracted from the recording.
This corresponds a summary statistics model.
For $M = 4$, most methods perform well, since this allows for better characterization of various signal structures in the scenes.
As the number of clusters increases, the \emph{RbQ-a} method performs significantly worse for both scattering features and MFCCs since any distinct objects are averaged out by clusters representing the background.
The \emph{RbQ-w} and \emph{RbQ-c} methods do better in this regard, as they are better able to emphasize the clusters that discriminate well between scenes.

Using \emph{RbQ-c}, therefore, we are not very sensitive to the choice of $M$, as long as it is large enough to allow for separation of the discriminative sound objects from the background.
This motivates our choice of $M = 8$ for the previous experiments in this section.

\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{figures/sensitivityM}
\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precision at rank $5$ ($p@5$) obtained for different methods as a function of the number of clusters $M$.}
\label{fig:sensitivityM}
\end{center}
\end{figure}

>>>>

The experiments in this paper are carried out on the DCASE 2013 dataset~\cite{7100934}.
Although the dataset was constructed for the task of acoustic scene classification, where the goal is to correctly assign the class of a given recording, we can use the same recordings and class labels for the task of acoustic scene similarity retrieval.
The dataset consists of two parts, a public and a private subset, each made up of $100$ acoustic scene recordings sampled at $44100\,\mathrm{Hz}$ and $30$ seconds in duration. The dataset is evenly divided into $10$ acoustic scene classes: \emph{bus}, \emph{busy street}, \emph{office}, \emph{open air market}, \emph{park}, \emph{quiet street}, \emph{restaurant}, \emph{supermarket}, \emph{tube}, and \emph{tube station}.
The recordings were made by three different recordists at a wide variety of locations in the Greater London area over a period of several months.
In order to avoid any correlation between recording conditions and label distribution, all recordings were carried out under moderate weather conditions, at varying times of day and week, and each recordist recorded each scene type. As a result, the dataset enjoys significant intra-class diversity while remaining of manageable size, making it suitable for evaluation of algorithmic design choices~\cite{lagrange:hal-01082501}.
As an illustration, Figure \ref{fig:scalogram} represents the wavelet scalogram of one recording within the DCASE 2013, labeled as \emph{park}.

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figures/park_scalogram.png}
\caption{
\label{fig:scalogram}
Wavelet scalogram $\boldsymbol{x}_1(t,\gamma_1)$ of the audio recording \emph{park04} in the DCASE 2013 dataset. We observe that this acoustic scene is a mixture of transient events (chirping birds, footsteps) and stationary texture (flowing water).}
\end{center}
\end{figure}

@@

The scaling parameter $q$ of the RBF kernels (see Eq.~\ref{eq:kc}) is set to $10\%$ of the number of data points to cluster.  As the number of Gaussians for the BoF approach, the numbers of clusters $M$ controls the level of abstraction. For each method, the parameter $M$ is set to $8$. It thus allows $8$ different types of observations to de modeled, which seems reasonable given the duration of the scene (30 seconds).

@@
